\documentclass[12pt,letterpaper, onecolumn]{exam}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage[lmargin=71pt, tmargin=1.2in]{geometry}
\thispagestyle{empty}

\begin{document}

\section*{Solutions}
\begin{enumerate}

    \item \textbf{Determine the phasor of the sinusoidal signal $f(t) = 10\sin{(\frac{\pi}{4}t - \frac{3\pi}{4})}$? Express it both in polar and rectangular.}

        To solve this question, let's first transform the sine function to a cosine function. To do so, we can subtract $\frac{\pi}{2}$, or $90\degree$, from the phase:
    
        \[
            f(t) = 10\sin{\left(\frac{\pi}{4}t - \frac{3\pi}{4}\right)} =
            10\cos{\left(\frac{\pi}{4}t - \frac{3\pi}{4} - \frac{\pi}{2}\right)}
            = 10\cos{\left(\frac{\pi}{4}t - \frac{5\pi}{4}\right)}
        \]
    
        From this form, we get that the polar form of this function is $10e^{-\frac{5\pi}{4}j}$. Expressing this in rectangular form, we have:
        \[
            10 \left( \frac{-\sqrt{2}}{2} \right) + 10j \left( \frac{\sqrt{2}}{2} \right) = -5\sqrt{2} + 5\sqrt{2}j
        \]

    \item \textbf{Determine the phasor of the sinusoidal signal $f(t) = 4\cos{(5t + \frac{\pi}{4})} + 12\cos{(t + \frac{7\pi}{4})}$? Express it either polar or rectangular form.}

        To solve this, we must first note that the two terms have \textbf{different frequencies} ($\omega_1 = 5$ rad/s and $\omega_2 = 1$ rad/s). Phasors are frequency-specific representations; therefore, we cannot add these two terms into a single phasor representation. We must determine the phasor for each component separately.

        \textbf{Component 1 ($\omega = 5$):}
        \[
            \mathbf{F_1} = 4\angle{\frac{\pi}{4}} = 4e^{j\pi/4}
        \]
        Rectangular: 
        \[
            4\left(\cos\frac{\pi}{4} + j\sin\frac{\pi}{4}\right) = 2\sqrt{2} + j2\sqrt{2}
        \]

        \textbf{Component 2 ($\omega = 1$):}
        \[
            \mathbf{F_2} = 12\angle{\frac{7\pi}{4}} = 12e^{j7\pi/4}
        \]
        Rectangular: 
        \[
            12\left(\cos\frac{7\pi}{4} + j\sin\frac{7\pi}{4}\right) = 6\sqrt{2} - j6\sqrt{2}
        \]
        
        Thus, the signal is represented by the superposition of these two distinct phasors.

    \item \textbf{Given an LTI system with a frequency response $H(\omega) = \frac{2}{12 + j\omega}$ and an input of $f(t) = 7 + \sin{(5t + \frac{\pi}{2})} + \cos{(12t + \frac{\pi}{4})}$, determine the output $y(t)$.}

        To solve this problem, we use the property of superposition and determine the output for each frequency component separately.

        \textbf{Part 1 (DC, $\omega=0$):}
        The input is $7$ (which is $7\cos(0t)$).
        \[
             H(0) = \frac{2}{12 + j(0)} = \frac{1}{6}
        \]
        The output is the input magnitude times the DC gain:
        \[
             y_1(t) = 7 \cdot H(0) = 7 \cdot \frac{1}{6} = \frac{7}{6}
        \]

        \textbf{Part 2 ($\omega=5$):}
        The input is $\sin{(5t + \frac{\pi}{2})} = \cos(5t)$.
        Evaluate frequency response at $\omega=5$:
        \[
            H(5)= \frac{2}{12 + 5j}
        \]
        Magnitude:
        \[
            |H(5)| = \frac{2}{\sqrt{12^2 + 5^2}} = \frac{2}{\sqrt{144+25}} = \frac{2}{13}
        \]
        Phase:
        \[
            \angle H(5) = -\arctan{\left(\frac{5}{12}\right)}
        \]
        Output component:
        \[
            y_2(t) = \frac{2}{13}\cos\left(5t - \arctan{\left(\frac{5}{12}\right)}\right)
        \]

        \textbf{Part 3 ($\omega=12$):}
        The input is $\cos{(12t + \frac{\pi}{4})}$.
        Evaluate frequency response at $\omega=12$:
        \[
            H(12) = \frac{2}{12 + 12j} = \frac{1}{6(1+j)}
        \]
        Magnitude:
        \[
             |H(12)| = \frac{1}{6\sqrt{1^2+1^2}} = \frac{1}{6\sqrt{2}} = \frac{\sqrt{2}}{12}
        \]
        Phase of H(12):
        \[
            \angle H(12) = -\arctan{\left(\frac{12}{12}\right)} = -\frac{\pi}{4}
        \]
        Total phase = Input phase + System phase = $\frac{\pi}{4} - \frac{\pi}{4} = 0$.
        Output component:
        \[
            y_3(t) = \frac{\sqrt{2}}{12}\cos(12t)
        \]

        \textbf{Total Output:}
        Combining all parts:
        \[
            y(t) = \frac{7}{6} + \frac{2}{13}\cos{\left(5t - \arctan{\left(\frac{5}{12}\right)}\right)} + \frac{\sqrt{2}}{12}\cos(12t)
        \]

    \item \textbf{Consider the RC circuit with $R = 2\ \Omega$, $C = 1\ \mathrm{F}$, and input $f(t)=4e^{-t}u(t)$ with $v(0^-)=1$.}

        To solve this problem, we first write KCL at the capacitor node. The current through the resistor plus the current into the capacitor must sum to zero:

        \[
            \frac{v(t)-f(t)}{R} + C\frac{dv(t)}{dt} = 0
        \]

        Plugging in $R=2$ and $C=1$:

        \[
            \frac{v(t)-4e^{-t}}{2} + \frac{dv(t)}{dt} = 0
        \]

        Multiplying by 2 gives the differential equation:

        \[
            2\frac{dv(t)}{dt} + v(t) = 4e^{-t}
        \]

        Solving this ODE, we rewrite it as:

        \[
            \frac{dv(t)}{dt} + \frac{1}{2}v(t) = 2e^{-t}
        \]

        Using the integrating factor $e^{t/2}$, we get:

        \[
            \frac{d}{dt}\left(v(t)e^{t/2}\right) = 2e^{-t/2}
        \]

        Integrating both sides:

        \[
            v(t)e^{t/2} = -4e^{-t/2} + K
        \]

        Solving for $v(t)$:

        \[
            v(t) = -4e^{-t} + Ke^{-t/2}
        \]

        Applying the initial condition $v(0)=1$:

        \[
            1 = -4 + K \quad \Rightarrow \quad K=5
        \]

        Thus, the full solution is:

        \[
            v(t) = -4e^{-t} + 5e^{-t/2}
        \]

\item \textbf{Consider the following periodic signal $f(t)$.}

\begin{questions}
\question Given the fundamental frequency of $f(t)$ is 3 rad/s and $f(t)$ has exponential Fourier series coefficients:
    \[F_n = \frac{jn}{1+n+n^2}\]

Determine the output if $f(t)$ is the input to an LTI system with frequency response:
\begin{center}
    $H(\omega) = 4e^{-j2\omega}$ if $\omega \in [-8,8]$ and zero else
\end{center}

\textbf{Solution: } First notice that the frequency response $H(\omega)$ only has non-zero outputs for frequencies between $\pm 8$ rad/s. The Fourier series works with multiples of the fundamental frequency $\omega_0$, so the only Fourier series coefficients that produce a non-zero output are up to $F_{\pm 2}$. Notice that $n=3$ has a frequency of $3 \cdot 3 = 9$ rad/s, which is outside the range $[-8,8]$. \\

The output Fourier series coefficients $Y_n$ are given by $Y_n = \sum_{n=-2}^{2} F_n \cdot H(n\omega_0)$. Start with $n=0$.

\[
F_0 = 0,\ H(0) = 4
\]
Since $F_0 = 0$, there is no DC term in the output. Now we can calculate the rest of the Fourier series coefficients and their corresponding frequency response.

\begin{align*}
F_1 = \frac{j}{3},\ H(3) = 4e^{-j9} \\
F_{-1} = \frac{-j}{1},\ H(-3) = 4e^{j9} \\
F_2 = \frac{j2}{7},\ H(6) = 4e^{-j18} \\
F_{-2} = \frac{-j2}{3},\ H(-6) = 4e^{j18}
\end{align*}

Now we have the Fourier series coefficients for the output. Recall that this is an exponential Fourier series with fundamental frequency $\omega_0 = 3$ rad/s, so each coefficient is multiplied with $e^{jn\omega_0t}$ to produce the full output. Hence,

\[
\boxed{y(t) = \frac{j4}{3}e^{-j9}e^{j3t} - j4e^{j9}e^{-j3t} + \frac{j8}{7}e^{-j18}e^{j6t}- \frac{j8}{3}e^{j18}e^{-j6t}}
\]
This solution could be simplified further, but this form is sufficient to earn full points so it is recommended to save time by not simplifying if not needed.

\newpage


\question The signal $f(t)$ goes through another LTI system with output Fourier series coefficients:
\begin{center}
    $G_1 = j,\ G_{-1} = -j,\ G_2 = j\frac{4}{3},\ G_{-2} = -j\frac{4}{3},\ G_3 = j\frac{1}{2},\ G_{-3} = -j\frac{1}{2}$, and zero else
\end{center}
Determine the average power of the output signal. \\

\textbf{Solution: } Recall Parseval's Theorem from your formula sheet: $P = \sum_{n=-\infty}^{\infty}|F_n|^2$. In this case $n \in [-3,3]$, so our average power will be given by $P = \sum_{n=-3}^{3}|G_n|^2$. The magnitude of a completely imaginary component is just the coefficient next to $j$, so the average power is simply: 

\[
P_G = (1)^2+(-1)^2+(\frac{4}{3})^2+(-\frac{4}{3})^2+(\frac{1}{2})^2+(-\frac{1}{2})^2
\]

Simplifying, we get that 

\[
P_G = 1+1+\frac{16}{9}+\frac{16}{9}+\frac{1}{4}+\frac{1}{4} = \boxed{\frac{109}{18}}
\]

\end{questions}

\item \textbf{Consider the function $f(t) = \cos{(6t)} + 3\sin{(18t-\frac{\pi}{2})} - e^{j(21t + \frac{3\pi}{4})}$.}

\begin{questions} 
\question Is $f(t)$ periodic? \\

\textbf{Solution: } Yes, $f(t)$ is periodic. For a sum of periodic functions to be periodic itself, the ratio of every frequency has to be a rational number. That is, $\frac{\omega_i}{\omega_j} \in \mathbb{Q}$ where $i \neq j$ for every combination of frequencies. \\

In this function, we can see that $\omega_1 = 6,\ \omega_2 = 18,\ \omega_3 = 21$. 

\begin{align*}
    \frac{\omega_1}{\omega_2} = \frac{6}{18} \in \mathbb{Q} \\
    \frac{\omega_1}{\omega_3} = \frac{6}{21} \in \mathbb{Q} \\
    \frac{\omega_2}{\omega_3} = \frac{18}{21} \in \mathbb{Q}
\end{align*}

All possible combinations of frequencies are rational numbers, so $f(t)$ is periodic.

\newpage

\question If $f(t)$ is periodic, determine its period and fundamental frequency. If not, explain why. \\

\textbf{Solution: } From part 1, we already found that $f(t)$ is periodic. The fundamental frequency of the sum of periodic functions is given by gcd($\omega_1,\omega_2,\omega_3,...$). Therefore,

\[
\omega_0 = \gcd(6,18,21) = \boxed{3\ [rad/s]}
\]

Notice that 6, 18, and 21 are all divisible by 3 and no larger number can divide all three. \\

The period T of a periodic function is simply $\frac{1}{f_0}$ where $f_0$ is the fundamental frequency. However, in this class we mainly work with angular frequencies, so we need to use the relationship $\omega_0 = 2\pi f_0$ to find the standard fundamental frequency. 

\[
f_0 = \frac{\omega_0}{2\pi} = \frac{3}{2\pi}
\]

Therefore, T = $\frac{1}{f_0} = \boxed{\frac{2\pi}{3}\ [s]}$

\end{questions}

\end{enumerate}

\end{document}