\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{mathtools}

\title{ECE 310 Midterm 1 Worksheet}
\author{By HKN Members}
\date{}

\begin{document}

\maketitle

\section{Ethan's Remix}
Pramod has written a song, and wants Ethan to make a remix of it. To do this, Ethan creates a system described by the following LCCDE:

$ \frac{1}{4} y[n-2] - \frac{4}{9} y[n] = \frac{4}{3} x[n-1] + \frac{8}{9} x[n]$

After running it on the song, is this system guaranteed to not cause Pramod's song to become unbounded?

\section{V's MP}
V has measured that his productivity on his ECE 411 MP is described by the system $\frac{1}{n} x[n]$, where $n$ is the number of days until the MP is due, and $x[n]$ is how much work he accomplishes as a function of the number of days since the MP was assigned. Would delaying V's starting day have an effect on how much work he will accomplish in $k$ days?

\section{Proving LTIC (is fake)}
Given the $5$ systems below, prove or disprove that each one is LTIC.
\newline
\newline
\indent
$1. \ y[n]=\sin(2024n)$
\newline
\newline
\indent
$2. \ y[n]-y[n-2]=x[n-1]+x[n+2]$
\newline
\newline
\indent
$3. \ y[n]=\cos(x[n])$
\newline
\newline
\indent
$4. \ y[n]= n^2x[n] + x[n-2]$
\newline
\newline
\indent
$5. \ y[n]=\frac{1}{5}x[|n|]$

\section{Insider Trading?}

Shomik plays golf with Ethan after Ethan worked at a trading company. Ethan tells Shomik to construct a system over Ethan's company stock average every day as a way of calculating how much to buy of it. This system is described by the impulse response:

$\{9,4,\underset{\uparrow}{2},0,2,4,6\}$


It is illegal to make trades based in not publicly known (future) information. After implementing this system successfully with Ethan's help, will the SEC arrest them?

\section{Convoluted Convolving}
Use your favorite method of convolution to convolve these two signals:
\newline
\indent \indent \indent \indent 
$x[n]=\{\underset{\uparrow}{2},0,2,4\}$ \indent \indent $h[n]=\{-1,\underset{\uparrow}{3},2,0\}$

\section{Catching These Z-transforms}

\subsection{Inverse Z-Transforms}
Take the Inverse Z-transform of these signals and state the ROC. Determine whether each system is BIBO stable or not as well.
\newline\newline
$\frac{1}{1-z^{-1}}$, Right Sided
\newline\newline
$\frac{1}{1-z^{-1}}$, Left Sided
\newline\newline
$\frac{1 + z^{-1}}{z^{-2} + z^{-1} + \frac{1}{4}}$, Right Sided

\subsection{Z-Transform}

Find the Z-Transform of the following systems in terms of $X(z)$:
\newline\newline
$\frac{n}{4} x[n-4] $
\newline\newline
$\frac{3}{4} x[n+5] + x[n]$
\newline\newline
Find the Z-Transform of the following equations:
\newline\newline
$n^2 u[n-1] + (\frac{1}{2})^n u[n+1]$
\newline\newline
$u[n+1] * u[n-1]$


\end{document}
