\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{mathtools}

\title{HKN ECE 310 Review Worksheet 2 Solutions}
\author{Author: Members of HKN }
\date{}
\usepackage[makeroom]{cancel}
\usepackage[letterpaper, portrait, margin=1in]{geometry}
\pagenumbering{arabic}

\begin{document}

\maketitle

\section{The very basics}
\begin{enumerate}
    \item What is the relation between the Z-transform and the Discrete-Time Fourier Transform? When is this relation not valid?
    \textcolor{red}{$$ X_{d}(\omega) = X(z)\vert_{z=\exp(j\omega)} $$}

\end{enumerate}

\section{Devious DTFTs}
Use your unbounded knowledge of the DTFT properties and pairs to take the DTFT of these signals.
\newline\newline
(i) $\delta[n-1] + (\frac{2}{3})u[n+4]$
\newline\newline
(ii) $\cos(n)\sin(n)$
\newline\newline
(iii) $n\delta[n-1]*e^{jn}$

\

\noindent \textbf{Solution}: 

\

\noindent (i) We can apply the table and DTFT properties to get:

\textcolor{red}{$$\delta[n-1] + \frac{2}{3}u[n+4] \xleftrightarrow{\mathcal{F}} e^{-jw} + \frac{2}{3}e^{j4w}(\frac{1}{1-e^{-jw}} + \pi \delta(w))$$}

\noindent (ii) We could take the DTFT of both $\cos(n)$ and $\sin(n)$ and then convolve them in the frequency domain, but we opt to choose a simpler solution. Recall the folloing trig identity:

\textcolor{red}{$$\sin(\alpha)\cos(\beta) = \frac{1}{2}(\sin(\alpha + \beta) + \sin(\alpha-\beta))$$}

\noindent Applying it to our problem, we see that:

\textcolor{red}{$$\sin(n)\cos(n) = \frac{1}{2}(\sin(n + n) + \sin(n-n)) = \frac{1}{2}\sin(2n)$$}

\noindent The problem is now quite straightforward and we can just apply the table:

\textcolor{red}{$$\sin(n)\cos(n) \xleftrightarrow{\mathcal{F}} \frac{\pi}{2j} (\delta(w-2) - \delta(w+2))$$}

\noindent (iii) Recall that convolution in the time domain is multiplication in the frequency domain. We first find the DTFT of both input functions:

\textcolor{red}{$$n\delta[n-1] \xleftrightarrow{\mathcal{F}} j \frac{d}{dw}(e^{-jw}) = j(-je^{-jw}) = e^{-jw}$$
$$e^{jn} \xleftrightarrow{\mathcal{F}} 2\pi \delta(w-1)$$}

\noindent Now it's a straightforward frequency domain multiplication:

\textcolor{red}{$$n\delta[n-1]*e^{jn} \xleftrightarrow{\mathcal{F}} 2\pi e^{-jw} \delta(w-1)$$}

\noindent (iii) \textbf{Alternate Solution} A slightly quicker way to arrive at the same result is to notice that $n\delta[n-1] = \delta[n]$. It's easy to take the convolution of a function with a delta function, so $n\delta[n-1]*e^{jn} = e^{j(n-1)}$. It's straightforward to apply DTFT properties from here to get the same answer as above.

\section{Sampling and DTFTs}
(Let the output of a radio be) Consider a signal given by
$$
x(t) = 2\cos(10\pi t) + \sin(30\pi t)
$$
\begin{enumerate}
    \item What is the nyquist sampling rate of this signal?
    
    \textcolor{red}{Since the maximum frequency of any of the components of this signal is 15Hz, the minimum sampling rate must be $2\cdot 15$Hz $= 30$Hz.}

    \item Let's say the signal is sampled at twice the nyquist rate. What does the discrete-time signal look like for three samples starting at $n=0$? What is the $n$'th sample?

    \textcolor{red}{We are sampling at $60$Hz, so the sampled signal is
    $$x[n] = 2\cos\left( \frac{1}{6}\pi n \right) + \sin\left( \frac{1}{2}\pi n \right)$$}
   
    \item Find the Discrete-Time Fourier Transform of the signal. Plot both the real and imaginary components of the DTFT over the range $(-\pi, \pi)$.

    \textcolor{red}{The DTFT is
    $$X_{d}(\omega) = 2\pi \left[ \delta\left(\omega - \frac{\pi}{6}\right) + \delta\left(\omega + \frac{\pi}{6}\right) \right] - j \pi \left[ \delta\left(\omega - \frac{\pi}{2}\right) - \delta\left(\omega + \frac{\pi}{2}\right) \right]$$}

    \item What is the power contained in this signal? Make sure to include units!

    \textcolor{red}{Using Parseval's Relation, the power is 5$\pi$ watts.}

    \item If we want to build a low-pass filter to filter out the fastest component of this signal, what is the smallest value of $\omega$ at which the filter can start attenuating? 

    \textcolor{red}{$$\omega_\text{cutoff} = \frac{\pi}{6}$$}

    \item Lets say we have a perfect filter to do said filtering. We then amplify the signal such that the magnitude of each component is doubled. How does the power of the signal change?

    \textcolor{red}{Using Parseval's relation again, the power contained is $16\pi$ watts.}
    
\end{enumerate}

\end{document}
